---
title: "Apple's music world"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyverse)
source("compmus.R")


data <- read.csv("/Users/wuke/Desktop/apple/compmus2025.csv")


data$filename <- as.character(data$filename)


song1_name <- "ke-w-1"
song2_name <- "ke-w-2"


my_songs <- data %>% filter(filename %in% c(song1_name, song2_name))


class_avg <- data %>% summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))



head(my_songs)
```

## Welcome to Apple's music world!!!
Welcome to my music world, where I will use what I learned in my computational musicology class to analyze two songs I made with AI.

```{r}

```


### Introduction
🎵 Song 1: Melodic Summer Vibes

The first piece was created with prompts focusing on melodic summer vibes and energetic bass, running for 2 minutes and 30 seconds. 


```{r}

```



🎵 Song 2: Deep House Energy

The second track combines elements of J-pop and R&B with groovy rhythms and dreamy melodies, featuring a clear verse-chorus structure and reverb effects, also lasting 2 minutes and 30 seconds.

```{r}

```

### Music visualization

#### Part A
```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)


# calculate the average valuse for class
class_avg <- data %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) %>%
  mutate(filename = "Class Average")

# combine my songs with the class average data
final_data <- bind_rows(my_songs, class_avg)

# convert data to long format for visualization
final_data_long <- final_data %>%
  pivot_longer(cols = -filename, names_to = "feature", values_to = "value")


# all the features
data_no_tempo <- final_data_long %>% filter(feature != "tempo")

# tempo
data_tempo <- final_data_long %>% filter(feature == "tempo")

# creat bar charts
# all the features
bar_plot_no_tempo <- ggplot(data_no_tempo, aes(x = feature, y = value, fill = filename)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Comparison of Features (Excluding Tempo)", x = "Feature", y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# tempo
bar_plot_tempo <- ggplot(data_tempo, aes(x = feature, y = value, fill = filename)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Comparison of Tempo", x = "Feature", y = "Value")

# combine charts
interactive_bar_combined <- subplot(
  ggplotly(bar_plot_no_tempo),
  ggplotly(bar_plot_tempo),
  nrows = 1  
)

# final visualization
interactive_bar_combined



```

#### Part B

```{r}
"features/ke-w-1.json" |>                           
  compmus_chroma(norm = "identity") |>                 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(option = "plasma") +               
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()    
```

#### PartC

```{r}
"features/ke-w-1.json" |>                           
  compmus_mfccs(norm = "identity") |>                  
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      

```

#### Part D
```{r}
library(cowplot)
chroma_ssm <-"features/ke-w-1.json" |>                           
  compmus_chroma(norm = "identity") |>                 
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  ggtitle("Chroma Self-Similarity") +
  theme_classic() +
  theme(
    text = element_text(size = 14),  
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")  
  ) +
  coord_fixed()


mfcc_ssm <- "features/ke-w-1.json" |>                   
  compmus_mfccs(norm = "identity") |>                  
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  ggtitle("MFCC Self-Similarity") +
  theme_classic() +
  theme(
    text = element_text(size = 14),  
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")  
  ) +
  coord_fixed()

plot_grid(chroma_ssm, mfcc_ssm, ncol = 2, align = "hv", rel_widths = c(1, 1))

```

### What do we get from my music？？？
My two AI-generated songs stand out from the class average, particularly in terms of tempo, arousal, and valence. Essentia has effectively captured the characteristics of my music, clearly distinguishing my tracks from the class average in these aspects. Overall, my songs share similarities with other tracks in the class corpus, but they also exhibit unique features, such as a faster tempo and stronger emotional intensity.
```{r}
```
